---
title: "Generating a balance sheet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating a balance sheet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is an example usage to rebuild the balance sheet of a company's most recent
filing. This may not work for all filings. I will show where the common breaks
are and how it can be corrected.

# Setup

You will need to load the XBRL package found at
[github:jbennett7/XBRL](https://github.com/jbennett7/XBRL)

```{r setup}
# Load secfilings (this package)
devtools::load_all('.')
library(dplyr)
library(XBRL)
library(purrr)
library(tidyr)
```

You will need to set a user agent in order to download files from SEC EDGAR.

```{r useragent}
useragent <- "Your Name yourname@example.com"
```

# Downloading the filing

We will download the most recent filing of Google from EDGAR. (`slice(1)` works
because `get_submissions` returns the rows in reverse-chronological order which
is how EDGAR's JSON is ordered.)  EDGAR uses accession numbers to uniquely
identify SGML documents. See `filing_urls` in `R/get_filing.R` for the
downloading URL. When extracting the data from the SGML file, multiple files are
generated inside a temporary directory. Ensure that the directory is cleaned up
before extracting multiple filings. `xbrlDoAll` will generate a list of data
frames that will hold all the XBRL data. See the `XBRL` package for more
information.

```{r download-extract-filing}
cik <- get_cik('GOOG', useragent)

# We use the accessionNumber to retrieve the file see `get_filing`
accn <- get_submissions(cik, useragent) |>
    filter(grepl("^10-[KQ](?:\\/A)?$", form)) |>
    slice(1) |> pull(accessionNumber)

sgml <- get_filing(cik, accn, useragent)

# The file is extracted into multiple files placed in a temporary directory.
outdir <- extract_instance(sgml)

inst <- list.files(outdir,
    pattern="(?:\\d{8}|_htm)\\.xml", full.names=TRUE)

# The data is stored in multiple data frames in this list.
xbrl_data <- xbrlDoAll(inst)

# Clean up the output directory after generating the xbrl data list.
#unlink(outdir, recursive=TRUE)
```

## Locating the balance sheet role

This is where the code may break for different companies. It all depends on what
the company calls the "balance sheet." Most companies will call it
consolidated balance sheets. If you search the `xbrl_data$role` data frame for
`type == 'Statement'`. You will see what the company calls each of its financial
statements. The `ignore.case` option is used because sometimes a company will
label it something like: `CONSOLIDATED BALANCE SHEETS` or `Consolidated balance
sheets` or `Consolidated Balance Sheets`.

```{r get-role-id}
balance_sheet_role_id <- xbrl_data$role |>
    filter(
        type == 'Statement',
        grepl('balance sheets?$', ignore.case=TRUE, description)) |>
    pull(roleId)
```

## Building the presentation tree

Here we tap several data frames within the xbrl data. We need parent/child
relationships between each of the facts, the order of the facts and what their
preferred label is. Please review XBRL concepts such as `hypercubes` and
`dimensions` for more information about how XBRL organizes and represents
multidimensional facts. (We will be using dimensionless facts because we are
interested in company wide data. Please see explaination in the last section.)
Then we get the label strings and provide a function for applying the correct
label to the correct element.

```{r presentation-label}
pres <- xbrl_data$presentation |>
    filter(roleId == balance_sheet_role_id) |>
    select(fromElementId, toElementId, order, preferredLabel) |>
    mutate(order = as.numeric(order))

elements <- xbrl_data$element |>
    select(elementId, type, substitutionGroup, abstract) |>
    mutate(structural = substitutionGroup %in%
        c("xbrldt:hypercubeItem", "xbrldt:dimensionItem") |
        type == "dtr-types:domainItemType")

labels <- xbrl_data$label |>
    select(elementId, labelRole, labelString)

label_for <- function(elt, preferred_role) {
    cand <- labels |> filter(elementId == elt)
    hit <- cand |> filter(labelRole == preferred_role)
    if (nrow(hit) == 0) hit <- cand |> filter(labelRole == "http://www.xbrl.org/2003/role/label")
    if (nrow(hit) == 0) return(elt)
    hit$labelString[1]
}
```

## Flattening the tree into statement order

We define two more functions for searching through and flattening the
presentation tree. Another fragile point is that `root` assumes exactly one
root. If a filer's presentation linkbase has multiple disconnected top-level
nodes, `root` becomes a vector and `flatten()` isn't written to hand that.

```{r dfs-over-tree}
flatten <- function(id, depth, preferred_role, seq_env) {
    is_struct <- elements$structural[match(id, elements$elementId)]
    is_struct <- isTRUE(is_struct)
    row <- NULL
    if (!is_struct) {
        seq_env$n <- seq_env$n + 1
        row <- tibble(seq = seq_env$n, elementId = id, depth = depth,
                       negated = grepl("negated", preferred_role %||% ""),
                       label = label_for(id, preferred_role %||%
                           "http://www.xbrl.org/2003/role/label"))
    }
    next_depth <- if (is_struct) depth else depth + 1
    kids <- pres |> filter(fromElementId == id) |> arrange(order)
    child_rows <- pmap(list(kids$toElementId, kids$preferredLabel),
                        \(cid, plabel) flatten(cid, next_depth, plabel, seq_env))
    bind_rows(row, bind_rows(child_rows))
}

`%||%` <- function(a, b) if (is.null(a) || is.na(a)) b else a

root <- setdiff(pres$fromElementId, pres$toElementId)
seq_env <- new.env()
seq_env$n <- 0L
statement <- flatten(root, depth = 0, preferred_role = NA, seq_env = seq_env)
statement <- statement |> filter(elementId != root) # drop the [Abstract] doc-title node
```

## Extracting the values

 We need to ensure we get the correct facts. Some facts are part of company
 segments. For example "APAC" for Asian operations. These facts will carry the
 same name, but will have a dimension to them. We need to ensure we get company
 wide values, not segmented values. In practice, this means the values are
 "dimensionless". We also need to ensure the facts are instant facts, not
 duration facts. Balance Sheets are point in time documents.

```{r pull-data}
elt_ids <- unique(statement$elementId)
values <- xbrl_data$fact |>
    filter(elementId %in% elt_ids) |>
    inner_join(xbrl_data$context |> filter(instant), by="contextId") |>
    anti_join(xbrl_data$dimension, by="contextId") |>
    distinct(elementId, endDate, fact, decimals) |>
    mutate(fact = as.numeric(fact),
           precision = if_else(decimals == "INF", Inf, suppressWarnings(as.numeric(decimals)))) |>
    group_by(elementId, endDate) |>
    slice_max(precision, n = 1, with_ties = FALSE) |>
    ungroup() |>
    select(-decimals, -precision)

date_cols <- head(sort(unique(values$endDate), decreasing = TRUE), 2)
values <- values |> filter(endDate %in% date_cols)

wide <- values |>
    pivot_wider(names_from = endDate, values_from = fact) |>
    select(elementId, all_of(date_cols))

balance_sheet <- statement |>
    left_join(wide, by = "elementId") |>
    arrange(seq) |>
    mutate(across(all_of(date_cols), ~ if_else(negated, -.x, .x))) |>
    mutate(display_label = paste0(strrep("  ", depth), label)) |>
    select(display_label, all_of(date_cols))

print(balance_sheet, n = Inf)
```
