---
title: "statement_analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{statement_analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(secfilings)
library(XBRL)
library(dplyr)

# Need when accessing data from EDGAR.
useragent <- "Your Name yourname@example.com"
```

A good use of `secfilngs` and `XBRL` is to analyze financial statements. In the
`balance_sheet.Rmd` vignette we constructed a balance sheet for Google. Here we
will take specific facts and do ratio analysis with Google and similar cloud
competitors: Amazon and Microsoft.

This will be very crude at first. I just want to show how secfilings and XBRL
are used together. More sophisticated functions can be created using these
primitives.

First some helper functions

```{r helper-functions}
# Only pull the latest submission
get_latest_xbrl <- function(ticker, useragent) {
    cik <- secfilings::get_cik(ticker, useragent)
    
    accn <- secfilings::get_submissions(cik, useragent) |>
        filter(grepl("^10-[KQ](?:\\/A)?$", form)) |>
        slice(1) |> pull(accessionNumber)

    sgml <- secfilings::get_filing(cik, accn, useragent)
    outdir <- secfilings::extract_instance(sgml)
    
    inst <- list.files(outdir,
        pattern="(?:\\d{8}|_htm)\\.xml$", full.names=TRUE)
    
    xbrl <- XBRL::xbrlDoAll(inst)
    
    unlink(outdir, recursive=TRUE)
    return(xbrl)
}

# Extract an instant fact that matches element_id
get_instant_fact_value <- function(xbrl, element_id) {
    xbrl$fact |>
        filter(elementId == element_id) |>
        inner_join(xbrl$context |> filter(instant), by = "contextId") |>
        anti_join(xbrl$dimension, by = "contextId") |>
        mutate(fact = as.numeric(fact)) |>
        filter(endDate == max(endDate)) |>
        slice(1) |>
        pull(fact)
}

# First ratio: the current ratio
current_ratio <- function(xbrl) {
    current_assets <- get_instant_fact_value(xbrl, "us-gaap_AssetsCurrent")
    current_liabilities <- get_instant_fact_value(xbrl, "us-gaap_LiabilitiesCurrent")
    current_assets / current_liabilities
}

# Second ratio: total debt ratio
total_debt_ratio <- function(xbrl) {
    total_assets <- get_instant_fact_value(xbrl, "us-gaap_Assets")
    total_equity <- get_instant_fact_value(xbrl,
        "us-gaap_StockholdersEquity")
    print(total_equity)
    total_debt_ratio <- (total_assets - total_equity) / total_assets
}
```

```{r download-latest-quarter}
goog_xbrl <- get_latest_xbrl('GOOG', useragent)
goog_cr <- current_ratio(goog_xbrl)
goog_tdr <- total_debt_ratio(goog_xbrl)
amzn_xbrl <- get_latest_xbrl('AMZN', useragent)
amzn_cr <- current_ratio(amzn_xbrl)
amzn_tdr <- total_debt_ratio(amzn_xbrl)
msft_xbrl <- get_latest_xbrl('MSFT', useragent)
msft_cr <- current_ratio(msft_xbrl)
msft_tdr <- total_debt_ratio(msft_xbrl)
c(goog_cr, amzn_cr, msft_cr)
c(goog_tdr, amzn_tdr, msft_tdr)
```
